---
title: "Challenge 9"
author: "Austin Liu"
format: html
editor: visual
---

## Data Import and Packages

```{r}
#| message: false
library(tidyverse)
library(DT)

names <- read_csv(here::here("supporting_artifacts", 
                             "learning_targets",  
                             "Lab 9", 
                             "StateNames_A.csv"))
```

```{r}
datatable(names)
```

# Part 1: Summarizing and Visualizing Allisons

## Question 1

```{r}
names <- names |> 
  mutate(Sex = Gender)

allisonname <- names |> 
  select(Name, Year, State, Count, Sex) |> 
  group_by(State, Sex) |> 
  filter(Name == "Allison") |> 
  summarize(Count = sum(Count), .groups = "keep") |> 
  pivot_wider(names_from = Sex, values_from = Count) |> 
  mutate(M = coalesce(M, 0))

knitr::kable(allisonname[, 1:3], 
             col.names = c("State", 
                           "Total Sum of Female", 
                           "Total Sum of Male" ),
             "html")
```

## Question 2

```{r}
allisonname_F <- names |> 
  filter(Name == "Allison", Sex == "F")

```

## Question 3

```{r}
allisonname_f_byYear <- allisonname_F |> 
  group_by(Year) |> 
  summarize(Count = sum(Count))

  ggplot(data = allisonname_f_byYear, mapping = aes(x = Year, y = Count)) +
  geom_point() +
    geom_line() +
  labs(title = 'Popularity of the name "Allison" over time')

```

# Part 2: Modeling the Number of Allisons

## Question 4

```{r}
Model1 <- allisonname_f_byYear |> 
  lm(Count ~ Year, data = _)
```

## Question 5

```{r}
Model1 |> 
  ggplot(mapping = aes(y = Count, x = Year)) +
  geom_point() +
  stat_smooth(method = "lm")

```

## Question 6

```{r}
lm(Count ~ Year, data = allisonname_f_byYear)

```

y-hat (estimated Count of Allisons) = 209,689.8 - 101.5 (Year)

## Question 7

```{r}
Model1 |> 
  broom::augment() |> 
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```

In the plot of the residuals against the fitted values, we do not see any discernible pattern.

## Question 8

Our model shows that the name Allison is declining in popularity. Allison is still quite popular as about 5000 newborn babies were given that name in our most recent year of data.

# Part 3: Spelling by State

## Question 1

```{r}
#| message: false
names |> 
  filter(Sex == "M", Name %in% c("Allan", "Alan", "Allen")) |> 
  group_by(Year, Name) |> 
  summarize(Count = sum(Count)) |> 
  ggplot(mapping = aes(x = Year, y = Count, color = Name)) +
    geom_point() +
    geom_line() +
  labs(title = 'Popularity of the name "Allen, Allan, Alan" over time')
```

## Question 2

```{r}
alan_name_M <- names |> 
  filter(Sex == "M", Name %in% c("Allan", "Alan", "Allen"),
         Year == 2000, State %in% c("PA", "CA")) |> 
  pivot_wider(names_from = Name, values_from = Count) |> 
  select(State, Alan, Allen, Allan)
alan_name_M

knitr::kable(alan_name_M[, 1:4], 
             col.names = c("State",
                           "Count of Alan", 
                           "Count of Allen", 
                           "Count of Allan"),
             "html")
```

## Question 3

```{r}
alan_name_M_per <- names |> 
  filter(Sex == "M", Name %in% c("Allan", "Alan", "Allen"),
         Year == 2000, State %in% c("PA", "CA")) |> 
  group_by(State) |> 
  mutate(Count = Count/sum(Count)) |> 
  pivot_wider(names_from = Name, values_from = Count) |> 
  select(State, Alan, Allen, Allan)
alan_name_M_per
  
knitr::kable(alan_name_M_per[, 1:4], 
             col.names = c("State", 
                           "Percent by State named Alan", 
                           "Percent by State named Allen", 
                           "Percent by State named Allan") , 
             "html") %>%
  kableExtra::kable_styling(latex_options = "striped", font_size = 13)%>%
  kableExtra::row_spec(1:2, color = 'white', background = 'black')
```
