---
title: "Lab 4"
author: "Austin Liu"
format: html
editor: visual
---

## Data Import and Packages

```{r}
#| message: false
library(tidyverse)
library(here)
library(tidyr)

avocado <- read_csv(here("supporting_artifacts", 
                         "learning_targets", 
                         "Lab 4", 
                         "avocado.csv")) |> 
  rename("SmallAvo" = `4046`,
         "LargeAvo" = `4225`,
         "XLargeAvo" = `4770`)
```

# Question 1

This data set represents weekly retail data for Hass Avocados sold from 2015 to 2018. This data set includes the date, average price of a single avocado, the type of avocado, year, city or region, total volume of avocados sold, and total number of avocados sold by product look up number.

# Question 2

```{r}
#| message: false
#Question 2
region <- data.frame(region = 
              c("West",
               "Northeast",
               "SouthCentral",
               "Southeast", 
               "GreatLakes", 
               "Plains",
               "Midsouth",
               "WestTexNewMexico"))

regions <- semi_join(avocado, region)

state <- data.frame(region = c("California", "NewYork", "SouthCarolina"))

states <- semi_join(avocado, state)

metro <- avocado |>
  filter(region != "TotalUS") |> 
  anti_join(regions) |> 
  anti_join(states)
```

# Question 3

```{r}
#Question 3
q3 <- regions |> 
  group_by(region) |> 
  filter(type == "organic", year == 2017) |> 
  select(region, year, SmallAvo) |>
  summarize(across(.cols = everything(), mean, na.rm = TRUE))
q3
slice_max(q3, order_by = SmallAvo)
```

The Westregion sold the most small, organic Hass avocados on average in 2017.

# Question 4

```{r}
#Question 4
q4 <- avocado |> 
  separate(Date, into = c("Year", "Month", "Day")) |> 
  group_by(Month) |> 
  select(Month, `Total Volume`) |>
  summarize(across(.cols = everything(), mean, na.rm = TRUE))
q4
slice_max(q4, order_by = `Total Volume`)
```

February was the month with the highest total average volume of avocado sales.

# Question 5

```{r}
#Question 5
top5 <- metro |> 
  group_by(region) |> 
  select(`Total Volume`, region) |> 
  summarize(across(.cols = everything(), .fns = mean)) |> 
  slice_max(order_by = `Total Volume`, n = 5)

top5box <- metro |> 
  semi_join(top5, by = "region")

ggplot(data = top5box, mapping = aes(x = `Total Volume`, y = region)) + 
  geom_boxplot()
```

# Question 6

```{r}
#| message: false
#Question 6
cali <- data.frame(region = c("Sacramento", "LosAngeles", "SanDiego", "SanFrancisco"))

california <- semi_join(avocado, cali)

differences <- california |> 
  pivot_wider(names_from = region, values_from = AveragePrice) |>  
  group_by(type) |> 
  select(type, LosAngeles, SanFrancisco, 
         Sacramento, SanDiego) |> 
  summarize(across(.cols = everything(), .fns = mean, na.rm = TRUE)) |> 
  pivot_longer("LosAngeles":"SanDiego", names_to = "Cali_Region", values_to = "Avg Price")

differences
```

```{r}
ggplot(data = differences, mapping = aes(x = `Avg Price`, y = Cali_Region, color = type)) + 
  geom_point() +
  labs(title = "Comparison of Avg Price of Avocados by type",
       y = "Region of California", x = "Average Price ($)") +
   annotate("text", y = 1, x = 1.1, label = "Conventional") +
  annotate("text", y = 2, x = 1.1, label = "Conventional") +
  annotate("text", y = 3, x = 1.25, label = "Conventional") +
  annotate("text", y = 4, x = 1.2, label = "Conventional") +
   annotate("text", y = 1, x = 1.6, label = "Organic") +
  annotate("text", y = 2, x = 2.1, label = "Organic") +
  annotate("text", y = 3, x = 1.9, label = "Organic") +
  annotate("text", y = 4, x = 2, label = "Organic") + 
  theme(legend.position = "none")
  
```

# Question 7

```{r}
#Question 7
proportions <- california |> 
  pivot_longer(SmallAvo:XLargeAvo, names_to = "Avocado Size", values_to = "NumAvoSold") |> 
  group_by(region, type) |> 
  mutate(Frequency = NumAvoSold / sum(NumAvoSold))
  

ggplot(data = proportions, mapping = aes(x = region, y = Frequency, fill = `Avocado Size`)) +
  geom_col() + 
  facet_wrap(~type) + 
  labs(x = "Region of CA", y = "Proportion of Mean Avocadoes Sold")
  
```
